---
title: "PS1"
author: "Gage Clawson"
date: "1/12/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(dplyr)
library(readxl)
library(RColorBrewer)
```

```{r}
adults <- read_excel("04_Data_set_1.xlsx", sheet = "adults")
recruits <- read_excel("04_Data_set_1.xlsx", sheet = "recruits")
```

##Question 1
###Compare and contrast changes in mean density for adults 
Make a histogram of mean denisty for each organism (y-axis) on both Rarotonga and Aitutaki in 2009 and 2018
  
```{r}
surgeon <- adults %>% 
  filter(organism == "surgeonfish") 
yellow <- adults %>%
  filter(organism == "yellow damsel")
spotted <- adults %>%
  filter(organism == "spotted damsel")
trout <- adults %>%
  filter(organism == "coral trout")
sea_an <- adults %>%
  filter(organism == "sea anemones")
mound_coral <- adults %>%
  filter(organism == "mounding corals")
fine_coral <- adults %>%
  filter(organism == "fine-branching corals")
coarse_corals <- adults %>%
  filter(organism == "coarse-branching corals")
algae <- adults %>%
  filter(organism == "algae")
```

```{r}
##surgeon graph
ggplot(surgeon, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.95, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
   scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Adult Surgeonfish Mean Density")

##spotted damsel graph
ggplot(spotted, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Adult Spotted Damsel Mean Density")

##yellow damsel graph
ggplot(yellow, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Adult Yellow Damsel Mean Density")

##trout graph
ggplot(trout, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Adult Coral Trout Mean Density")

##sea anemones graph
ggplot(sea_an, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Sea Anemones Mean Density")

##mounding corals graph
ggplot(mound_coral, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Mounding Corals Mean Density")

##fine-branching corals graph
ggplot(fine_coral, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Fine-branching Corals Mean Density")


## coarse-branching corals corals graph
ggplot(coarse_corals, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Coarse-branching Corals Mean Density")

## algae graph
ggplot(algae, aes(x = location, y = density, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Mean Density (per 100 m^2)",
       x = "Location",
       title = "Algae Mean Density")

```


##Question 2
###Compare and contrast changes in mean density for recruits 
Make a histogram of mean denisty for each organism (y-axis) on both Rarotonga and Aitutaki in 2009 and 2018

```{r}
surgeon_recruit <- recruits %>% 
  filter(fish == "surgeonfish") 
yellow_recruits <- recruits %>%
  filter(fish == "yellow damsel")
spotted_recruits <- recruits %>%
  filter(fish == "spotted damsel")
trout_recruits <- recruits %>%
  filter(fish == "coral trout")

```

```{r}
##surgeon graph
ggplot(surgeon_recruit, aes(x = location, y = recruits, fill = location)) +
  geom_col(width = 0.95, alpha = 0.7) +
  geom_errorbar(aes(ymin = recruits - ci, ymax = recruits + ci), width = 0.5) +
  facet_wrap(~year) +
   #scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Recruit Mean Density (per m^2 of Fine-branching Coral)",
       x = "Location",
       title = "Recruit Surgeonfish Mean Density")

##spotted damsel graph
ggplot(spotted_recruits, aes(x = location, y = recruits, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = recruits - ci, ymax = recruits + ci), width = 0.5) +
  facet_wrap(~year) +
   #scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Recruit Mean Density (per m^2 of Sea Anenome)",
       x = "Location",
       title = "Recruit Spotted Damsel Mean Density")

##yellow damsel graph
ggplot(yellow_recruits, aes(x = location, y = recruits, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = recruits - ci, ymax = recruits + ci), width = 0.5) +
  facet_wrap(~year) +
   #scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Recruit Mean Density (per m^2 of Mounding Coral)",
       x = "Location",
       title = "Recruit Yellow Damsel Mean Density")

##trout graph
ggplot(trout_recruits, aes(x = location, y = recruits, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  geom_errorbar(aes(ymin = recruits - ci, ymax = recruits + ci), width = 0.5) +
  facet_wrap(~year) +
   #scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Recruit Mean Density (per m^2 of Coarse-branching Coral)",
       x = "Location",
       title = "Recruit Coral Trout Mean Density")

```

###Question 3
Calculate Recruits per 100 m2 reef area and graph adult vs recruits scatterplots

```{r}
recruits_q3 <- read_excel("04_Data_set_1.xlsx", sheet = "recruits_3") %>%
  filter(recruits_100m > 0) 
  #left_join(adults, by = c("fish" = "organism"))

surgeon_recruit_q3 <- recruits_q3 %>% 
  filter(fish == "surgeonfish") 
yellow_recruits_q3 <- recruits_q3 %>%
  filter(fish == "yellow damsel")
spotted_recruits_q3 <- recruits_q3 %>%
  filter(fish == "spotted damsel")
trout_recruits_q3 <- recruits_q3 %>%
  filter(fish == "coral trout")

plot(surgeon$density, surgeon_recruit_q3$recruits_100m, xlab = "Adult Density (per 100m2)", ylab = "Recruitment Density (per 100m2)", main = "Surgeonfish Adult vs Recruitment Density")
abline(fit1 <- lm(surgeon_recruit_q3$recruits_100m ~ surgeon$density))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit1)$r.squared, digits=4)))

plot(yellow$density, yellow_recruits_q3$recruits_100m, xlab = "Adult Density (per 100m2)", ylab = "Recruitment Density (per 100m2)", main = "Yellow Damsel Adult vs Recruitment Density")
abline(fit2 <- lm(yellow_recruits_q3$recruits_100m ~ yellow$density))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit2)$r.squared, digits=4)))

plot(spotted$density, spotted_recruits_q3$recruits_100m, xlab = "Adult Density (per 100m2)", ylab = "Recruitment Density (per 100m2)", main = "Spotted Damsel Adult vs Recruitment Density")
abline(fit3 <- lm(spotted_recruits_q3$recruits_100m ~ spotted$density))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit3)$r.squared, digits=4)))

plot(trout$density, trout_recruits_q3$recruits_100m, xlab = "Adult Density (per 100m2)", ylab = "Recruitment Density (per 100m2)", main = "Coral Trout Damsel Adult vs Recruitment Density")
abline(fit4 <- lm(trout_recruits_q3$recruits_100m ~ trout$density))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit4)$r.squared, digits=4)))
```

###Question 4
```{r}
plot(fine_coral$density, surgeon$density, xlab = "Fine-branching Coral Density (m2 per 100m2)", ylab = "Surgeonfish Adult Density (per 100m2)", main = "Fine-branching Coral Density vs Surgeonfish Adult Density")
abline(fit_fine <- lm(surgeon$density ~ fine_coral$density ))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit_fine)$r.squared, digits=4)))

plot(coarse_corals$density, trout$density, xlab = "Coarse-branching Coral Density (m2 per 100m2)", ylab = "Coral Trout Adult Density (per 100m2)", main = "Coarse-branching Coral Density vs Coral Trout Adult Density")
abline(fit_coarse <- lm(trout$density ~ coarse_corals$density ))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit_coarse)$r.squared, digits=4)))

plot(mound_coral$density, yellow$density, xlab = "Mounding Coral Density (m2 per 100m2)", ylab = "Yellow Damsel Adult Density (per 100m2)", main = "Mounding Coral Density vs Yellow Damsel Adult Density")
abline(fit_mound <- lm(yellow$density ~ mound_coral$density ))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit_mound)$r.squared, digits=4)))

plot(sea_an$density, spotted$density, xlab = "Sea Anenome Density (m2 per 100m2)", ylab = "Spotted Damsel Adult Density (per 100m2)", main = "Sea Anenome Density vs Spotted Damsel Adult Density")
abline(fit_sea <- lm(spotted$density ~ sea_an$density ))
legend("top", bty="n", legend=paste("R2 =", 
       format(summary(fit_sea)$r.squared, digits=4)))
```

###Question 5
```{r}
adults_5 <- read_excel("04_Data_set_1.xlsx", sheet = "adults_5")
```

```{r}
surgeon_5 <- adults_5 %>% 
  filter(organism == "surgeonfish") 
yellow_5 <- adults_5 %>%
  filter(organism == "yellow damsel")
spotted_5 <- adults_5 %>%
  filter(organism == "spotted damsel")
trout_5 <- adults_5 %>%
  filter(organism == "coral trout")

```

```{r}
##surgeon graph
ggplot(surgeon_5, aes(x = location, y = density_per_micro, fill = location)) +
  geom_col(width = 0.95, alpha = 0.7) +
  #geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
   #scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0), limits = c(0,150)) +
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Density (per 100 m^2 of Fine-branching Coral)",
       x = "Location",
       title = "Adult Surgeonfish Density per Area Fine-branching Coral Habitat")

##spotted damsel graph
ggplot(spotted_5, aes(x = location, y = density_per_micro, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  #geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
   scale_y_continuous(expand = c(0,0), limits = c(0,525)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Density (per 100 m^2 of Sea Anemome)",
       x = "Location",
       title = "Adult Spotted Damsel Density per Area Sea Anemome Habitat")

##yellow damsel graph
ggplot(yellow_5, aes(x = location, y = density_per_micro, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  #geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
   scale_y_continuous(expand = c(0,0), limits = c(0,5.5)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Density (per 100 m^2 of Mounding Coral)",
       x = "Location",
       title = "Adult Yellow Damsel Density per Area Mounding Coral Habitat")

##trout graph
ggplot(trout_5, aes(x = location, y = density_per_micro, fill = location)) +
  geom_col(width = 0.90, alpha = 0.7) +
  #geom_errorbar(aes(ymin = density - ci, ymax = density + ci), width = 0.5) +
  facet_wrap(~year) +
   #scale_y_continuous(expand = c(0,0), limits = c(-20,450), breaks = seq(0,400,100)) + 
  scale_x_discrete(expand = c(0,0)) + 
  scale_y_continuous(expand = c(0,0), limits = c(0,55)) +
  scale_fill_manual(name = "", values = c("grey51", "royalblue")) +
  theme_bw() + 
  theme(legend.position = "none") +
  labs(y = "Adult Density (per 100 m^2 of Coarse-branching Coral)",
       x = "Location",
       title = "Adult Coral Trout Density per Area Coarse-branching Coral Habitat")
```

###Question 6
Make scatterplots of all Adult Densities

 - surgeonfish vs coral trout
 - surgeonfish vs yellow damsel
 - surgeonfish vs spotted damsel
 - coral trout vs spotted damsel
 - coral trout vs yellow damsel
 - yellow damsel vs spotted damsel

```{r}
library(ggpmisc)
#surgeon vs trout
surgtroutlm <- lm(trout$density~surgeon$density)

ggplot() + 
  geom_point(aes(x = surgeon$density, y = trout$density)) +
  geom_smooth(aes(y = trout$density, x = surgeon$density), method = lm, se = FALSE) +
  theme_bw() +
  labs(x = "Adult Surgeonfish Density (# per 100 m2)", y = "Adult Coral Trout Density (# per 100 m2)", title = "Adult Surgeonfish vs. Adult Coral Trout Density") +
    annotate("text", x = 50, y = 150, label = paste("R2 =", format(summary(surgtroutlm)$r.squared, digits=4)))

#surgeon vs yellow damsel
surgyellowlm <- lm(yellow$density ~ surgeon$density)

ggplot() + 
  geom_point(aes(x = surgeon$density, y = yellow$density)) +
  geom_smooth(aes(x = surgeon$density, y = yellow$density), method = lm, se = FALSE) +
  theme_bw() +
  labs(x = "Adult Surgeonfish Density (# per 100 m2)", y = "Adult Yellow Damselfish Density (# per 100 m2)", title = "Adult Surgeonfish vs. Adult Yellow Damselfish Density") +
   annotate("text", x = 300, y = 80, label = paste("R2 =", format(summary(surgyellowlm)$r.squared, digits=4)))

#surgeon vs spotted damsel
surgspotlm <- lm(spotted$density ~ surgeon$density)

ggplot() + 
  geom_point(aes(x = surgeon$density, y = spotted$density)) +
  geom_smooth(aes(x = surgeon$density, y = spotted$density), method = lm, se = FALSE) +
  theme_bw() +
  labs(x = "Adult Surgeonfish Density (# per 100 m2)", y = "Adult Spotted Damselfish Density (# per 100 m2)", title = "Adult Surgeonfish vs. Adult Spotted Damselfish Density") +
 annotate("text", x = 50, y = 750, label = paste("R2 =", format(summary(surgspotlm)$r.squared, digits=4)))

#trout vs spotted damsel
troutspotlm <- lm(spotted$density ~ trout$density)

ggplot() + 
  geom_point(aes(x = trout$density, y = spotted$density)) +
  geom_smooth(aes(x = trout$density, y = spotted$density), method = lm, se = FALSE) +
  theme_bw() +
  labs(x = "Adult Coral Trout (# per 100 m2)", y = "Adult Spotted Damselfish Density (# per 100 m2)", title = "Adult Coral Trout vs. Adult Spotted Damselfish Density") +
 annotate("text", x = 25, y = 750, label = paste("R2 =", format(summary(troutspotlm)$r.squared, digits=4)))

#trout vs yellow damsel
troutyellowlm <- lm(yellow$density ~ trout$density)

ggplot() + 
  geom_point(aes(x = trout$density, y = yellow$density)) +
  geom_smooth(aes(x = trout$density, y = yellow$density), method = lm, se = FALSE) +
  theme_bw() +
  labs(x = "Adult Coral Trout (# per 100 m2)", y = "Adult Yellow Damselfish Density (# per 100 m2)", title = "Adult Coral Trout vs. Adult Yellow Damselfish Density") +
   annotate("text", x = 125, y = 80, label = paste("R2 =", format(summary(troutyellowlm)$r.squared, digits=4)))

#yellow vs spotted damsel
yellowspotlm <- lm(spotted$density ~ yellow$density)

ggplot() + 
  geom_point(aes(x = yellow$density, y = spotted$density)) +
  geom_smooth(aes(x = yellow$density, y = spotted$density), method = lm, se = FALSE) +
  theme_bw() +
  labs(x = "Adult Yellow Damselfish (# per 100 m2)", y = "Adult Spotted Damselfish Density (# per 100 m2)", title = "Adult Yellow Damselfish vs. Adult Spotted Damselfish Density") +
   annotate("text", x = 75, y = 750, label = paste("R2 =", format(summary(yellowspotlm)$r.squared, digits=4)))
```





